---
title: "Big Mart Sales Prediction"
output: html_document
---

## Loading Dependencies

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(corrplot)
library(xgboost)
library(caret)
library(cowplot)
library(ranger)
library(outliers)
```

## Loading Datasets

```{r}
getwd()=='C:/Users/Iqbal/Documents/Rprog/big-mart-sales'
train = fread("train.csv") 
test = fread("test.csv ") 
submission = fread("ss.csv")
```

## Structure of Dataset

```{r}
dim(train)
dim(test)
dim(submission)
```

```{r}
"
str(train)
str(test)
str(submission)

summary(train)
summary(test)
summary(submission)

names(train)
names(test)
names(submission)

"
```

## Removing Outliers from Target Variable --- not gud
```{r}
"
train[ Item_Outlet_Sales > quantile(combi$Item_Outlet_Sales, 0.99, na.rm = TRUE), .N ]


#Remove unwanted rows in data.table
train <- train[ Item_Outlet_Sales < quantile(train$Item_Outlet_Sales, 0.99, na.rm = TRUE) ]
"
```

## Combining train and test
```{r}
#train[,"Outlet_Identifier"]

test[,Item_Outlet_Sales := NA] 
combi = rbind(train, test) # combining train and test datasets dim(combi)
dim(combi)

```


## Unique Outlets and Items

```{r}

"
combi$Outlet_Identifier <- as.factor(combi$Outlet_Identifier)
levels(combi$Outlet_Identifier)

combi$Item_Identifier <- as.factor(combi$Item_Identifier)
nlevels(combi$Item_Identifier)
"
print(head(unique(combi$Item_Identifier)))
length(unique(combi$Item_Identifier))
is.factor(combi$Item_Identifier) 

print(unique(combi$Outlet_Identifier))
is.factor(combi$Outlet_Identifier) 
```

## Types of Items

```{r}
"
combi$Item_Type <- as.factor(combi$Item_Type)
levels(combi$Item_Type)

nlevels(combi$Item_Type) 
"
print(unique(combi$Item_Type))
is.factor(combi$Item_Type) 
```

## Types of Outlets

```{r}
print(unique(combi$Outlet_Type))
is.factor(combi$Outlet_Type)           
```

## Types of Outlet Location
```{r}
print(unique(combi$Outlet_Location_Type))
is.factor(combi$Outlet_Location_Type)
```

## Types of Outlet Size
```{r}
print(unique(combi$Outlet_Size))
is.factor(combi$Outlet_Size)
```

## Types of Fat Content

```{r}
print(unique(combi$Item_Fat_Content))
is.factor(combi$Item_Fat_Content)

```


## Feature Separation based on type of Data

  * Categorical Attributes - Freq dist table, bar charts, pie charts, paretto diagrams
  
    + Nominal
      - Item_Identifier
      - Item_Type
      - Outlet_Identifier
      
    + Ordinal
      - Item_Fat_Content
      - Outlet_Size
      - Outlet_Location_Type
      - Outlet_Type
      
      
  * Numerical Attributes
  
    + Discrete 
      - nlevels of Factor Variables
      - Outlet_Establishment_Year
      
    + Continuous
      - Item_Weight
      - Item_Visibility
      - Item_MRP
      - Item_Outlet_Sales

## Feature separation based on measurement levels
  
  * Qualitative -- similar to categorical data
    
  * Quantitative
  
    + Interval
      - Item_Weight
      - Item_Visibility
      - Item_MRP
      - Item_Outlet_Sales
      
    + Ratios
      - none

## Removing Dulpicates from Fat Content

```{r}
"
combi[Item_Fat_Content== low fat,Item_Fat_Content:=Low Fat]
combi[Item_Fat_Content== LF,Item_Fat_Content:=Low Fat]
combi[combi$Item_Fat_Content==reg,Item_Fat_Content:=Regular]

combi[combi$Item_Fat_Content== low fat]:=Low Fat
combi[combi$Item_Fat_Content==LF]:=Low Fat
combi[combi$Item_Fat_Content==reg]:=Regular
"
combi$Item_Fat_Content[combi$Item_Fat_Content == "LF"] = "Low Fat"
combi$Item_Fat_Content[combi$Item_Fat_Content == "low fat"] = "Low Fat"
combi$Item_Fat_Content[combi$Item_Fat_Content == "reg"] = "Regular"
##sample <- combi %>% group_by(Item_Fat_Content,Item_Type) %>% summarise(Count = n(), Item_mean_MRP = mean(Item_MRP))


ggplot(combi %>% group_by(Item_Fat_Content) %>% summarise(Count = n())) +   geom_bar(aes(Item_Fat_Content, Count), stat = "identity", fill = "coral1")
```


## Univariate Analysis

```{r}
ggplot(combi, aes(Outlet_Type,Outlet_Size,fill=Outlet_Type)) +
    geom_bar(stat = "identity", position= position_dodge())


```

```{r}

ggplot(combi) + 
    geom_histogram(aes(combi$Item_Outlet_Sales), binwidth=100, color= "black", fill="orange") +
    xlab("Item Outlet Sales")


```
```{r}
ggplot(combi) + 
    geom_histogram(aes(combi$Item_MRP), binwidth=1, color= "black", fill="orange") +
    xlab("Item MRP")
ggplot(combi) + 
    geom_histogram(aes(combi$Item_Weight), binwidth=0.05, color= "black", fill="orange") +
    xlab("Item Weight")
ggplot(combi) + 
    geom_histogram(aes(combi$Item_Visibility), binwidth=0.005, color= "black", fill="orange") +
    xlab("Item Visibility")


```



```{r}
ggplot(combi %>% dplyr::group_by(Item_Fat_Content) %>% dplyr::summarise(Count = n())) +   
      geom_bar(aes(Item_Fat_Content, Count), stat = "identity", fill = "blue")

ggplot(combi %>% dplyr::group_by(Item_Identifier) %>% dplyr::summarise(Count = n()) %>% arrange(desc(Count))) +
      geom_bar(aes(Item_Identifier, Count), stat = "identity", fill = "blue") +
      geom_label(aes(Item_Identifier, Count, label = Count), vjust = 0.5)+
      theme(axis.text.x = element_text(angle = 90, hjust=1)) + xlab("Item Identifier")

df <- combi %>% dplyr::group_by(Item_Type) %>% dplyr::summarise(Count = n())
ggplot(df) + 
      geom_bar(aes(Item_Type, Count), stat = "identity",position=position_dodge(), fill = "blue")+
      geom_label(aes(Item_Type, Count, label = Count), vjust = 0.5)+
      theme(axis.text.x = element_text(angle = 60, hjust=1)) + xlab("Item Type")

ggplot(combi %>% dplyr::group_by(Outlet_Type) %>% dplyr::summarise(Count = n())) + 
      geom_bar(aes(Outlet_Type, Count), stat = "identity", fill = "blue")+
      geom_label(aes(Outlet_Type, Count, label = Count), vjust = 0.5)

ggplot(combi %>% dplyr::group_by(Outlet_Identifier) %>% dplyr::summarise(Count = n())) +   
      geom_bar(aes(Outlet_Identifier, Count), stat = "identity", fill = "blue")+
      geom_label(aes(Outlet_Identifier, Count, label = Count), vjust = 0.5)

ggplot(combi %>% dplyr::group_by(Outlet_Size) %>% dplyr::summarise(Count = n()) %>% dplyr::arrange(desc(Count))) +   
      geom_bar(aes(Outlet_Size, Count), stat = "identity", fill = "blue")+
      geom_label(aes(Outlet_Size, Count, label = Count), vjust = 0.5)

ggplot(combi %>% dplyr::group_by(Outlet_Establishment_Year) %>% dplyr::summarise(Count = n())) + 
      geom_bar(aes(factor(Outlet_Establishment_Year), Count), stat = "identity", fill = "blue")+
      geom_label(aes(factor(Outlet_Establishment_Year), Count, label = Count), vjust = 0.5)+  
      xlab("Outlet_Establishment_Year") +  
      theme(axis.text.x = element_text(size = 8.5))


```

## Imputing Missing Values
```{r}
colSums(is.na(combi))
```

### Removing NA from Item_Weight

```{r}
missing_vals <- which(is.na(combi$Item_Weight))


for(i in missing_vals){
  item = combi$Item_Identifier[i]  
  combi$Item_Weight[i] = mean(combi$Item_Weight[combi$Item_Identifier == item], na.rm = T) }

```


### Removing NA from Item_Visibility

```{r}
missing_vals <- which(is.na(combi$Item_Visibility))


for(i in missing_vals){
  item = combi$Item_Identifier[i]  
  combi$Item_Visibility[i] = mean(combi$Item_Visibility[combi$Item_Identifier == item], na.rm = T) }

```

## Bivariate Analysis
```{r}
  # Item_Weight vs Item_Outlet_Sales 
ggplot(combi) +      
     geom_point(aes(Item_Weight, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +     
     theme(axis.title = element_text(size = 8.5))
# Item_Visibility vs Item_Outlet_Sales 
ggplot(combi) +       
      geom_point(aes(Item_Visibility, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +      
      theme(axis.title = element_text(size = 8.5))
# Item_MRP vs Item_Outlet_Sales 
ggplot(combi) +       
      geom_point(aes(Item_MRP, Item_Outlet_Sales), colour = "violet", alpha = 0.3) +      
      theme(axis.title = element_text(size = 8.5))

```

```{r}
# Item_Type vs Item_Outlet_Sales 
ggplot(combi) +       
  geom_violin(aes(Item_Type, Item_Outlet_Sales), fill = "magenta") +      
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.text = element_text(size = 6),axis.title = element_text(size= 8.5))
# Item_Fat_Content vs Item_Outlet_Sales 
ggplot(combi) + 
  geom_violin(aes(Item_Fat_Content, Item_Outlet_Sales), fill = "magenta") +      
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.text = element_text(size = 8),axis.title = element_text(size= 8.5))
# Outlet_Identifier vs Item_Outlet_Sales 
ggplot(combi) + 
  geom_violin(aes(Outlet_Identifier, Item_Outlet_Sales), fill = "magenta") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),axis.text = element_text(size = 8),axis.title = element_text(size= 8.5))

```

```{r}

ggplot(combi) + geom_violin(aes(Outlet_Location_Type, Item_Outlet_Sales), fill = "magenta") 
ggplot(combi) + geom_violin(aes(Outlet_Type, Item_Outlet_Sales), fill = "magenta") 


ggplot(combi) + geom_violin(aes(Outlet_Size, Item_Outlet_Sales), fill = "magenta")

```

## Feature Engineering

*Item_Type_new: Broader categories for the variable Item_Type.
*Item_category: Categorical variable derived from Item_Identifier.
*Outlet_Years: Years of operation for outlets.
*price_per_unit_wt: Item_MRP/Item_Weight
*Item_MRP_clusters: Binned feature for Item_MRP.

```{r}
perishable = c("Breads", "Breakfast", "Dairy", "Fruits and Vegetables", "Meat", "Seafood")
non_perishable = c("Baking Goods", "Canned", "Frozen Foods", "Hard Drinks", "Health and Hygiene", "Household", "Soft Drinks")

# create a new feature 'Item_Type_new' 
combi[,Item_Type_new := ifelse(Item_Type %in% perishable, "perishable", ifelse(Item_Type %in% non_perishable, "non_perishable", "not_sure"))]

```

Let’s compare Item_Type with the first 2 characters of Item_Identifier, i.e., ‘DR’, ‘FD’, and ‘NC’. These identifiers most probably stand for drinks, food, and non-consumable.

```{r}
table(combi$Item_Type, substr(combi$Item_Identifier, 1, 2))
combi[,Item_category := substr(combi$Item_Identifier, 1, 2)]
combi$Item_Fat_Content[combi$Item_category == "NC"] = "Non-Edible"
```

```{r}
# years of operation for outlets 
combi[,Outlet_Years := 2013 - Outlet_Establishment_Year] 
combi$Outlet_Establishment_Year = as.factor(combi$Outlet_Establishment_Year)

# Price per unit weight 
combi[,price_per_unit_wt := Item_MRP/Item_Weight]

combi[,Item_MRP_clusters := ifelse(Item_MRP < 69, "1st",ifelse(Item_MRP >= 69 & Item_MRP < 136, "2nd",ifelse(Item_MRP >= 136 & Item_MRP < 203, "3rd", "4th")))]
```


## Encoding Categorical Variables

```{r}
combi[,Outlet_Size_num := ifelse(Outlet_Size == "Small", 0,ifelse(Outlet_Size == "Medium", 1, 2))] 

combi[,Outlet_Location_Type_num := ifelse(Outlet_Location_Type == "Tier 3", 0,ifelse(Outlet_Location_Type == "Tier 2", 1, 2))]

combi[, c("Outlet_Size", "Outlet_Location_Type") := NULL]

```

```{r}

# One hot encoding for categorical variables
ohe = dummyVars("~.", data = combi[,-c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")], fullRank = T) 
ohe
summary(ohe)

```

```{r}
ohe_df = data.table(predict(ohe, combi[,-c("Item_Identifier", "Outlet_Establishment_Year", "Item_Type")])) 
```

```{r}
combi = cbind(combi[,"Item_Identifier"], ohe_df)
```

## Preprocessing

```{r}

#Skewed Variables

ggplot(combi) + 
    geom_histogram(aes(combi$price_per_unit_wt), binwidth=0.1, color= "black", fill="orange") +
    xlab("Item price_per_unit_wt")

ggplot(combi) + 
    geom_histogram(aes(combi$Item_Visibility), binwidth=0.005, color= "black", fill="orange") +
    xlab("Item Visibility")

ggplot(combi) + 
    geom_histogram(aes(combi$Item_Outlet_Sales), binwidth=0.05, color= "black", fill="orange") +
    xlab("Item Outlet Sales")


```

### Removing Skewness
```{r}
combi[,Item_Visibility := log(Item_Visibility + 1)] # log + 1 to avoid division by zero
combi[,price_per_unit_wt := log(price_per_unit_wt + 1)]
```
### Performing Log transform on Target Variable

```{r}
" TRIAL WORKS

train[,Item_Outlet_Sales_log_transform := log(Item_Outlet_Sales + 1)]
train[,Item_Outlet_Sales_exp_transform := exp(Item_Outlet_Sales_log_transform) - 1]

"

combi[,Item_Outlet_Sales := log(Item_Outlet_Sales + 1)]

```

### Normalization
```{r}
num_vars = which(sapply(combi, is.numeric)) # index of numeric features 
num_vars_names = names(num_vars) 
```

```{r}
combi_numeric = combi[,setdiff(num_vars_names, "Item_Outlet_Sales"), with = F] 


prep_num = preProcess(combi_numeric, method=c("center", "scale")) 
combi_numeric_norm = predict(prep_num, combi_numeric)


combi[,setdiff(num_vars_names, "Item_Outlet_Sales") := NULL] # removing numeric independent variables 
combi = cbind(combi, combi_numeric_norm)
```

```{r}

#Skewed Variables

ggplot(combi) + 
    geom_histogram(aes(combi$price_per_unit_wt), binwidth=0.1, color= "black", fill="orange") +
    xlab("Item price_per_unit_wt")

ggplot(combi) + 
    geom_histogram(aes(combi$Item_Visibility), binwidth=0.005, color= "black", fill="orange") +
    xlab("Item Visibility")

ggplot(combi) + 
    geom_histogram(aes(combi$Item_Outlet_Sales), binwidth=0.05, color= "black", fill="orange") +
    xlab("Item Outlet Sales")


```

# Data Splitting to Test and Train
```{r}
train = combi[1:nrow(train)] 
test = combi[(nrow(train) + 1):nrow(combi)] 
test[,Item_Outlet_Sales := NULL] # removing Item_Outlet_Sales as it contains only NA for test dataset

```

# Correlation
```{r}
cor_train = cor(train[,-c("Item_Identifier")]) 
corrplot(cor_train, method = "pie", type = "lower", tl.cex = 0.9)
```

## Model Building

```{r}
linear_reg_mod = lm(Item_Outlet_Sales ~ ., data = train[,-c("Item_Identifier")])
```

```{r}
dim(submission)
names(submission)
head(submission)
head(test)
```
```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(linear_reg_mod, test[,-c("Item_Identifier")])
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```

```{r}
# Define training control
set.seed(100) 
train.Control <- trainControl(method = "cv", number = 10)
# Train the model
model <- train(Item_Outlet_Sales ~ ., data = train[,-c("Item_Identifier")],
               method = "lm",
               trControl = train.Control)
```

```{r}
# Summarize the results
print(linear_reg_mod)
#plot(varImp(linear_reg_mod))
print(model)
```

```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(model, test[,-c("Item_Identifier")])
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```

## Lasso Regression

```{r}

set.seed(101) 
#my_control = trainControl(method="cv", number=5) 
Grid = expand.grid(alpha = 1, lambda = seq(0.001,0.1,by = 0.0002)) 
lasso_linear_reg_mod = train(x = train[, -c("Item_Identifier", "Item_Outlet_Sales")],
                             y = train$Item_Outlet_Sales,
                             method='glmnet', 
                             trControl= train.Control, 
                             tuneGrid = Grid)
```


```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(lasso_linear_reg_mod, test[,-c("Item_Identifier")])
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```

## Ridge Regression

```{r}
set.seed(102) 
#my_control = trainControl(method="cv", number=5) 
Grid = expand.grid(alpha = 0, lambda = seq(0.001,0.1,by = 0.0002)) 
ridge_linear_reg_mod = train(x = train[, -c("Item_Identifier", "Item_Outlet_Sales")],
                             y = train$Item_Outlet_Sales,
                             method='glmnet', 
                             trControl= train.Control,
                             tuneGrid = Grid)
```

```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(ridge_linear_reg_mod, test[,-c("Item_Identifier")])
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```

## Random Forrest

```{r}
set.seed(1237)
my_control = trainControl(method="cv", number=5) 
```

```{r}
tgrid = expand.grid(
  .mtry = c(3:10),
  .splitrule = "variance",
  .min.node.size = c(10,15,20)
)

rf_mod = train(x = train[, -c("Item_Identifier", "Item_Outlet_Sales")], 
               y = train$Item_Outlet_Sales,
               method='ranger', 
               trControl= my_control, 
               tuneGrid = tgrid,
               num.trees = 400,
               importance = "permutation")

```

```{r}
plot(rf_mod)
plot(varImp(rf_mod))
```

```{r}
# preparing dataframe for submission and writing it in a csv file 
submission$Item_Outlet_Sales = predict(rf_mod, test[,-c("Item_Identifier")])
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
write.csv(submission, "Linear_Reg_submit.csv", row.names = F)
```


## Xgboost Model

```{r}
param_list = list(
        objective = "reg:linear",
        eta=0.01,
        gamma = 1,
        max_depth=6,
        subsample=0.8,
        colsample_bytree=0.5
        )
```


```{r}
dtrain = xgb.DMatrix(data = as.matrix(train[,-c("Item_Identifier", "Item_Outlet_Sales")]),                          label= train$Item_Outlet_Sales)

dtest = xgb.DMatrix(data = as.matrix(test[,-c("Item_Identifier")]))
```

```{r}
set.seed(112)
 
xgbcv = xgb.cv(params = param_list, 
               data = dtrain, 
               nrounds = 1000, 
               nfold = 5, 
               print_every_n = 10, 
               early_stopping_rounds = 30, 
               maximize = F)
```

```{r}
xgb_model = xgb.train(data = dtrain, params = param_list, nrounds = 653)
```

```{r}
submission
submission$Item_Outlet_Sales <- exp(predict(xgb_model, data.matrix(test[,-c("Item_Identifier")])))-1
submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0
```

```{r}
write.csv(submission, "XGBoost_pred.csv", row.names = F)
```


## Variable Importance

```{r}
var_imp = xgb.importance(feature_names = setdiff(names(train), 
                                        c("Item_Identifier", "Item_Outlet_Sales")), 
                                        model = xgb_model)
xgb.plot.importance(var_imp)
```

## Weighted Average Ensembling

```{r}
linear <- predict(linear_reg_mod, test[,-c("Item_Identifier")])
K_fold <- predict(model, test[,-c("Item_Identifier")])
Lasso_reg <- predict(lasso_linear_reg_mod, test[,-c("Item_Identifier")])
Ridge_reg <- predict(ridge_linear_reg_mod, test[,-c("Item_Identifier")])
Random_Forest <- predict(rf_mod, test[,-c("Item_Identifier")])
XGBoost <- predict(xgb_model, data.matrix(test[,-c("Item_Identifier")]))

submission$Item_Outlet_Sales = (linear*0.05)+ (XGBoost*0.6) + (Random_Forest*0.25) + (Lasso_reg*0.05) + (Ridge_reg*0.05)

submission$Item_Outlet_Sales = exp(submission$Item_Outlet_Sales)-1

submission$Item_Outlet_Sales[submission$Item_Outlet_Sales<0]=0

write.csv(submission, "Weighted_Average.csv", row.names = F)
        
```
